---
title: "Are there qualitative individual differences in the SNARC effect?"
author: "Julia Haaf"
date: "9/11/2019"
bibliography      : "lab.bib"
output: pdf_document
header-includes:
   - \usepackage{bm}
   - \usepackage{pcl}
   - \usepackage{amsmath}
   - \usepackage{setspace}
csl: apa6.csl
---

\spacing{1.5}
\fontsize{12}{12}

- sanity check with 1+2 and 8+9
- Common-effect model include CHECK
- change to -1, 1 CHECK
- CIs for lm output CHECK

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(reshape2)
require(plyr)
require(QuantPsyc)
require(GeneNet)
require(Hmisc)
library(knitr)
library(BayesFactor)
library(MCMCpack)
library(plotrix)
library(RColorBrewer)

source("helpFunctions.R")

myCols <- brewer.pal(8, "Dark2")

meanScale <- 1/30
effectScale <- 1/30
```

I first looked at one data set with average proportions of negative, positive and null effects according to the preprint (Cipora et al, 2019). The data I picked are from Cipora (2014).

```{r}
# Loading the data
raw_data <- read.table("../data/Cipora_2014_raw_data.txt"
                       , sep="\t", dec=".", header = TRUE)

# Changing part.code variable to be a factor.
raw_data$part.code <- as.factor(raw_data$part.code)

## Julia data inspection
## I want data with accuracy and >200ms filter
# but not the 3SD filter they used.

dat <- subset(raw_data
              , correct.experimental == 1 & filter.correct.exp.antic == 1)
# around 30 trials per person and condition
mean(with(dat, tapply(rt, list(number, resp.side, part.code), length)))

#Needed later
prep <- prep.models(dat$part.code
                    , side = dat$resp.side
                    , number = dat$number)
```

## The Model

The model I used is a hierarchical model on the trial-level. Let $Y_{ijk\ell}$ be the response time for the $i$th person, the $j$th response side, the $k$the number, and the $\ell$th trial. Then 

\[
Y_{ijk\ell} \sim \mbox{Normal}(\mu + \alpha_i + x_j \nu_i + u_k \delta_i + v_{jk} \theta_i, \sigma^2),
\]

where $\mu$ is the grand mean, $\alpha_i$ is the $i$th person's baseline effect, $\nu_i$ is the $i$th person's response side effect, $\delta_i$ is the $i$th person's number effect, and $\theta_i$ is the $i$th person's SNARC effect, the congruency effect between side and number. The variable $x_j = \begin{cases}1,& \text{if } j = 1\\  0,& \text{if} j = 2\end{cases}$, the variable $u_k = k$, and the variable $v_{jk}$ is the product of $x_j$ and $u_k$.

## Observed SNARC

I use a similar linear model on each participant's data to get observed SNARC effects per person.

```{r}
# reorder participants
dat$sub <- prep$sub
dat <- dat[order(dat$sub), ]

# Averaging reaction times for each [participant x number x response side]
#configuration and forming the data frame in convenient manner, calculating dRTs
aggregated_data <- aggregate(rt~ resp.side + number + part.code, dat, mean)
aggregated_data_dRT <- dcast(aggregated_data, part.code + number ~ resp.side)
aggregated_data_dRT$dRT <- aggregated_data_dRT$right - aggregated_data_dRT$left

# Calculating unstandardized slopes with main effect of side and number
testSNARC <- ddply(aggregated_data, "part.code", function(x){
  regression <- lm(x$rt ~ x$number * x$resp.side, data = aggregated_data)
  ind.slope <- regression$coefficients[[4]]
  side.slope <- regression$coefficients[[2]]
  number.slope <- regression$coefficients[[3]]
  p.intercept <- regression$coefficients[[1]] 
  coef <- summary(regression)$coefficients[4,1]
  err <- summary(regression)$coefficients[4,2] 
  ci <- coef + c(-1,1) * err * qt(0.975, 12)
  data.frame(unstdSNARC.slope=ind.slope, side.slope, number.slope, p.intercept, lower = ci[1], upper = ci[2])
})
kable(head(testSNARC))
```

```{r, fig.asp = .7, fig.width=8}
cis <- testSNARC[, 6:7]
par(cex = 1.5, mgp = c(2, .7, 0))

plot(sort(testSNARC$unstdSNARC.slope), pch = "+", ylim = range(cis)
     , ylab = "SNARC effect (ms)"
     , xlab = "Participant"
     , col = 0
     , frame.plot = F
     , axes = F)
axis(1, c(1, prep$I))
axis(2, seq(-60, 30, 30))
abline(h = 0)
ind <- order(testSNARC$unstdSNARC.slope)
plotCI(x = 1: nrow(testSNARC)
       , y = testSNARC$unstdSNARC.slope[ind]
       , ui= cis[ind, 2]
       , li= cis[ind, 1]
       , add = TRUE
       , col = "gray40"
       , pch = 21
       , pt.bg= myCols[2]
       )
```

## Estimated SNARC

```{r estimation, cache = T}
out <- quid(id = dat$part.code
            , side = dat$resp.side
            , number = dat$number
            , prior = c(meanScale, effectScale)
            , rt = dat$rt, iter = 15000)
```

The estimated effects for the SNARC are flipped in sign for some reason. I will see if I can fix that, but for now I flip the observed SNARC effects for comparison. 

```{r, fig.asp = 1}
layout(matrix(1:4, ncol = 2, byrow = T))
par(mgp = c(2, .7, 0))
## check

#Individuals' effect of number
i.delta0 <- 1 * prep$I + 2 #index for the common effect
i.delta <- (1 * prep$I + 3):(2 * prep$I + 2) #indices for the individual effects

#add overall effect to individuals' deviations
myDelta <- out$mcmc.unconstrained[, i.delta] + 
  out$mcmc.unconstrained[, i.delta0]
plot(colMeans(myDelta), testSNARC$number.slope)
abline(0, 1)

#Individuals' effect of response side
i.nu0 <- 2 * prep$I + 3 #index for the common effect
i.nu <- (2 * prep$I + 4):(3 * prep$I + 3) #indices for the individual effects

#add overall effect to individuals' deviations
myNu <- out$mcmc.unconstrained[, i.nu] + 
  out$mcmc.unconstrained[, i.nu0] 
plot(colMeans(myNu), testSNARC$side.slope)
abline(0, 1)

#Individuals' baseline estimate
i.mu <- 1 #index for the common effect
i.alpha <- (2):(prep$I + 1) #indices for the individual effects

#add overall effect to individuals'
myAlpha <- out$mcmc.unconstrained[, i.mu] + 
  out$mcmc.unconstrained[, i.alpha]
plot(colMeans(myAlpha), testSNARC$p.intercept)
abline(0, 1)

#Individuals' SNARC effect
plot(out$ind.effects, testSNARC$unstdSNARC.slope)
abline(0, 1)
```

Some hierarchical shrinkage for the effects. I may change the scales for $\nu$ and $\delta$.

## Does anybody SNARC?

```{r, fig.asp = .7, fig.width=8}
cis <- apply(out$theta, 2, quantile, probs =  c(.025, .975))
par(cex = 1.5, mgp = c(2, .7, 0))

plot(sort(testSNARC$unstdSNARC.slope), pch = "+", ylim = c(-30, 30)
     , ylab = "SNARC effect (ms)"
     , xlab = "Participant"
     , col = myCols[2]
     , frame.plot = F
     , axes = F)
axis(1, c(1, prep$I))
axis(2, seq(-30, 30, 15))
abline(h = 0)
ind <- order(testSNARC$unstdSNARC.slope)
plotCI(x = 1: nrow(testSNARC)
       , y = out$ind.effects[ind]
       , ui= cis[2, ind]
       , li= cis[1, ind]
       , add = TRUE
       , col = "gray40"
       , pch = 21
       , pt.bg= myCols[1]
       )
```

Now. The estimates show quite a bit of hierarchical shrinkage. I might change the scales to see what happens. But They reflect the expected size of effect. We would expect effects around 5 ms on average, and the trial by trial variability to be around 150 ms. Likewise, we would expect variability around 5 ms. The scales therefore should be around 5 ms/150 ms, or $1/30$. So, the chosen scales are quite wide.

I first looked at the Bayes factor of no SNARC effect, the null model, and a varying SNARC effect, the unconstrained model. The Bayes factor is `r round(out$bfs[3], 2)` in favor of the null model.

```{r}
out$bfs
```

